Static Analysis by Path Focusing

### 2011-05-11 report/report.tex
removed all debug printing for performance evaluation + report update

\usepackage{mathcomp,amssymb,verimagstudent}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{mathenv}
\usepackage{tikz}
\usepackage{listings}
\usepackage{placeins}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

@@

	\begin{definition}[Liveness]
	A variable is \emph{live} at the control point $p$ iff:
	\begin{itemize}
	\item its value is available at $p$, meaning that there is a path from the
	point defining the variable and $p$.
	\item its value might be used in the future, meaning that there exist a path
	starting in $p$, that arrives in a point using the variable.
	\end{itemize}
	\end{definition}

@@

@@

@@

	To Be Completed

@@

\bibliographystyle{alpha}
  \bibliography{report}
\end{document}

>>>>

\usepackage{mathcomp,amssymb,verimagstudent}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{mathenv}
\usepackage{tikz}
\usepackage{listings}
\usepackage{placeins}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multirow}

@@

\subsection{Transformation passes}

LLVM internal representation can be applied transformation passes that modifies
parts of it. These passes are mostly used for optimisation purpose. In our
analyzer, we apply some already existing passes before running our 
abstract interpretation pass. Among the passes we run, one could cite:
\begin{itemize}
\item the \emph{PromoteMemoryToRegister} Pass, that transforms the internal
representation by promoting the memory variables to Static Single Assignment
(SSA) registers. This pass transform the control flow graph such that all the
variables are in SSA form, and removes most of the \emph{Load/Store} operations
into memory. It is useful since our analyzer doesn't have a memory model, and
then loses lots of precision in the case of \emph{Load/Store} instructions.
\item the \emph{LoopSimplify} and \emph{LoopInfo} Passes, that transform all loops into a canonical
format, and compute some information related to loops (for instance, the set of
BasicBlocks that are parts of the loop, etc.).
\item the \emph{LowerSwitch} Pass, that transform a switch instruction into its
corresponding \emph{if-then-else} instructions. After running this pass, there
is no more \emph{Switch} instruction in the control flow graph, so the
implementation of the static analysis is easier.
\end{itemize}

@@

	\begin{definition}[Liveness]
	A variable is \emph{live} at the control point $p$ iff:
	\begin{itemize}
	\item its value is available at $p$, meaning that there is a path from the
	point defining the variable to $p$.
	\item its value might be used in the future, meaning that there exist a path
	starting in $p$, that arrives in a point using the variable.
	\end{itemize}
	\end{definition}

@@

	In our running example (Figures \ref{runningexample},
	\ref{multigraphtransformGopan} and \ref{multigraphGopan}), the analyzer
	outputs the following result for the basicblock $1$:

\begin{C}
RESULT FOR BASICBLOCK: -------------------
; <label>:1                                 ; preds = %10, %0
  %x.0 = phi i32 [ 0, %0 ], [ %11, %10 ]
  %y.0 = phi i32 [ 0, %0 ], [ %y.1, %10 ]
  %2 = icmp sle i32 %x.0, 50
  br i1 %2, label %3, label %5
-----
environment: dim = (2,0), count = 77
 0: x.0
 1: y.0
polyhedron of dim (2,0)
array of constraints of size 3
 0: -x.0 - y.0 + 102 >= 0
 1: y.0 >= 0
 2: x.0 - y.0 >= 0
\end{C}

First, it shows the BasicBlock we work on, and its associated abstract value:
\begin{itemize}
\item The environment of the abstract value is its dimensions. Here, $x.0$ and
$y.0$ are the dimensions of the polyhedron.
\item The array of constraints defining the polyhedron. The three constraints
here define the polyhedra seen in Figure \ref{result}.b.
\end{itemize}

@@

\begin{tabular}{|c|c|c|c||c|c|} \hline

			& \multicolumn{3}{c||}{Path Focusing} & \multicolumn{2}{c|}{Lookahead
			Widening} \\ \hline
Benchmark	& \multicolumn{2}{c|}{time} & result	& time	& result	\\
\cline{2-3}	& z3 & Yices				& & & \\
\hline

@@

	In this report, we proposed a new approach to guide the iterations of the
	analyzer, in order to distinguish the different paths and to treat them
	independently. The loss of precision due to widening and control-flow merges
	is then reduced. This technique allows us, for instance, to do narrowing
	steps right after widening, or to use acceleration techniques when a path
	actually is a simple loop.

	We implemented this technique into a small analyzer, 
	and it appears that in lots of cases, 
	the precision of the result is improved 
	compared to the classical abstract interpretation technique.
	We compared it with \emph{Lookahead Widening}, and our technique seems to
	perform better in the case of loops containing control-flow merges.

\bibliographystyle{alpha}
  \bibliography{report}

@@

\begin{figure}
\begin{C}
(and (= bd_0 false)
        (= t_0_1 bs_0)
        (or (not bd_0) true)
        (= b_5 t_1_5)
        (= x_%6_ (+ x_y.0_ -1))
        (= t_5_7 b_5)
        (or (not b_5) true)
        (= b_3 t_1_3)
        (= x_%4_ (+ x_y.0_ 1))
        (= t_3_7 b_3)
        (or (not b_3) true)
        (= b_7
           (or t_5_7 t_3_7))
        (= t_7_9 (and b_7 (< x_y.1_ 0)))
        (= t_7_10
           (and b_7 (not (< x_y.1_ 0))))
        (or (not b_7)
            (= x_y.1_ (ite t_3_7 x_%4_ x_%6_)))
        (= b_10 t_7_10)
        (= x_%11_ (+ x_x.0_ 1))
        (= t_10_1 b_10)
        (or (not b_10) true)
        (= bd_1
           (or t_10_1 t_0_1))
        (= t_1_3 (and bs_1 (<= x_x.0_ 50)))
        (= t_1_5
           (and bs_1 (not (<= x_x.0_ 50))))
        (or (not bd_1)
            (and (= x'_x.0_ (ite t_0_1 0 x_%11_))
                 (= x'_y.0_ (ite t_0_1 0 x_y.1_))))
        (= b_9 t_7_9)
        (= t_9_12 b_9)
        (or (not b_9) true)
        (= bd_12 t_9_12)
        (or (not bd_12) true))
\end{C}
\caption{Expression of $\rho$ associated to the program in Figures
\ref{runningexample}, \ref{multigraphtransformGopan} and
\ref{multigraphGopan}.}
\end{figure}

\end{document}
